---
title: "Swiss dataset Exploration"
author: "Vetle Birkeland Huglen"
date: "18 9 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

In this part of the homework I will briefly explore the "swiss" dataset in R, which shows a handful of socio-economic indicators, along with fertility rate, for all French-speaking provinces of Switzerland. The statistics are from about 1888. Firstly, let us take a quick look at the format of the dataset:

```{r}
head(swiss)
```

It could also be useful to look at some summary statistics.

```{r}
summary(swiss)
str(swiss)
```
 Now, I wanna look further into the "fertility" feature.
 
```{r}
fertility = swiss$Fertility
mean(fertility)
max(fertility)
min(fertility)
sd(fertility)
var(fertility)
```
 
I wanna sort the provinces by fertilityrate, but I am only interested in the Fertility and Catholic features:

```{r}
rownames_to_column(swiss, var = "rowname") %>%
  select(rowname, Fertility, Catholic) %>%
    arrange(desc(Fertility))
```

Now, I notice that the provinces with highest Fertility also has quite high value of Catholic. Lets do a quick linear regression and see what we get!

```{r}
fit = lm(Fertility ~ Catholic, data=swiss)
summary(fit)
plot(swiss$Catholic, swiss$Fertility)
abline(fit)
```
This plot indicates a small rise in Fertility as function of Catholic. This is, however, a very bad linear regression. As seen in the plot, most of the variance in Fertility is left unexplained by the Catholic feature. Regardless, the low p-value shows that there with high probability exists some correlation between the two features.

This was a very shallow exploration of the swiss dataset, but it highlights some of the possibilities R possesses.
